The Galactic Midplane Is Not a Plane: Implications for Dynamical Analysis with \textit{Gaia} Data and Beyond

### 2019-05-05 paper/ms.tex
Add info about orbits in Fig 5

\begin{deluxetable*}{cccccccc}
\tablecaption{Description and names of the three orbits considered in this
work. For the last two columns: $z_{\text{max}}$ is the maximum height of the
orbit while $\frac{1}{2}(R_{\text{max}} - R_{\text{min}})$ is the magnitude of the radial excursions of the orbit. In the epicyclic approximation, $A_z = z_{\text{max}}$ and $A_R = \frac{1}{2}(R_{\text{max}} - R_{\text{min}})$ \label{tab:orbits}}
\tablehead{\colhead{name} & \colhead{initial position} & \colhead{initial
velocity} & \colhead{$J_R$} &
\colhead{$J_{\phi}$} & \colhead{$J_z$} & \colhead{$z_{\text{max}}$} &
\colhead{$\frac{1}{2}(R_{\text{max}} - R_{\text{min}})$}\\ \colhead{ } &
\colhead{$\mathrm{kpc}$} & \colhead{$\mathrm{km}/\mathrm{s}$} &
\colhead{$\mathrm{kpc\,km\,s^{-1}}$} &
\colhead{$\mathrm{kpc\,km\,s^{-1}}$} & \colhead{$\mathrm{kpc\,km\,s^{-1}}$} &
\colhead{$\mathrm{kpc}$} & \colhead{$\mathrm{kpc}$}}
\startdata 
thin-disk & $(8, 0, 0)$ & $(0, -190, 10)$ & 40.3 & -1520 & 0.69 & 0.12 & 1.29 \\
thick-disk & $(8, 0, 0)$ & $(0, -190, 50)$ & 32.5 & -1520 & 23.0 & 0.85 & 1.19 \\ 
halo & $(8, 0, 0)$ & $(0, -190, 190)$ & 32.8 & -1520 & 529.1 & 6.16 & 2.34
\enddata
\end{deluxetable*}

@@

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{fig/schmactions_many_orbits_Jz_fun.pdf}
\end{center}
\caption{The fractional error in $J_z$ as a function of $J_z$ for a few
different offsets in $z$. We show this for a $z$ offset of $10, 50, 100\,\pc$
(orange, teal, and green respectively). As before, the error ($\Delta J_z$) is one
half the 95\uth{} minus 5\uth{} percentile of the distribution of observed
$J_z$ values over the course of the orbit. There are large errors for
thin-disk like orbits ($J_z \sim 0.5\,\actunit$), even for a small midplane
offset of $10\,\pc$. }
\label{fig:dJz_fun_Jz}
\end{figure}

To further understand the effect of the midplane error, we also plot the
fractional error in $J_z$ as a function of $J_z$ for $z$ offsets of $10, 50,
100\,\pc$ (orange, teal, and green, respectively) in
Figure~\ref{fig:dJz_fun_Jz}. For a thin disk-like orbit
($J_z\sim0.5\,\actunit$), even a $10\,\pc$ offset in $z$ is enough to induce a
$\sim25\%$ error in $J_z$. For larger values of $J_z$, the fractional errors
are suppressed, but the induced error can still be large depending on how
great the $z$ offset is. We also plot the epicylic prediction for $\Delta J_z
/ J_z$ from Equation~\eqref{eq:Ji_err_mosttime} as dashed lines for each $z$
offset. We find that the epicyclic approximation matches the numerical
estimate quite well.

@@

\section{LSR Variations} \label{app:lsr}
We consider the variations of the LSR as a function of azimuth at the fiducial
solar circle ($R_{\odot} = 8.2\,\kpc$) in Figure~\ref{fig:lsr_variations}. At
each azimuth, $\phi$, we take the median velocity in cylindrical coordinates
of all stars within $200\,\pc$ of the position, following
\citet{2018arXiv180610564S}. No best-fit subtraction was performed as in
Figure~\ref{fig:midplane}.

>>>>

\begin{deluxetable*}{cccccccc}
\tablecaption{Description and names of the three orbits considered in this
work. For the last two columns: $z_{\text{max}}$ is the maximum height of the
orbit while $\frac{1}{2}(R_{\text{max}} - R_{\text{min}})$ is the magnitude of the radial excursions of the orbit. In the epicyclic approximation, $A_z = z_{\text{max}}$ and $A_R = \frac{1}{2}(R_{\text{max}} - R_{\text{min}})$. \label{tab:orbits}}
\tablehead{\colhead{name} & \colhead{initial position} & \colhead{initial
velocity} & \colhead{$J_R$} &
\colhead{$J_{\phi}$} & \colhead{$J_z$} & \colhead{$z_{\text{max}}$} &
\colhead{$\frac{1}{2}(R_{\text{max}} - R_{\text{min}})$}\\ \colhead{ } &
\colhead{$\mathrm{kpc}$} & \colhead{$\mathrm{km}/\mathrm{s}$} &
\colhead{$\mathrm{kpc\,km\,s^{-1}}$} &
\colhead{$\mathrm{kpc\,km\,s^{-1}}$} & \colhead{$\mathrm{kpc\,km\,s^{-1}}$} &
\colhead{$\mathrm{kpc}$} & \colhead{$\mathrm{kpc}$}}
\startdata 
thin-disk & $(8, 0, 0)$ & $(0, -190, 10)$ & 40.3 & -1520 & 0.69 & 0.12 & 1.29 \\
thick-disk & $(8, 0, 0)$ & $(0, -190, 50)$ & 32.5 & -1520 & 23.0 & 0.85 & 1.19 \\ 
halo & $(8, 0, 0)$ & $(0, -190, 190)$ & 32.8 & -1520 & 529.1 & 6.16 & 2.34
\enddata
\end{deluxetable*}

@@

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{fig/schmactions_many_orbits_Jz_fun.pdf}
\end{center}
\caption{The fractional error in $J_z$ as a function of $J_z$ for a few
different offsets in $z$. All orbits have the same initial position of $(8, 0, 0)\,\kpc$ and velocity $(-190, 0, v_z)\,\kms$, where we vary $v_z$. We show this for a $z$ offset of $10, 50, 100\,\pc$
(orange, teal, and green respectively). As before, the error ($\Delta J_z$) is
one half the 95\uth{} minus 5\uth{} percentile of the distribution of observed
$J_z$ values over the course of the orbit. There are large errors for
thin-disk like orbits ($J_z \sim 0.5\,\actunit$), even for a small midplane
offset of $10\,\pc$. As dashed lines in each color we also plot the prediction
for $\Delta J_z/J_z$ from the epicyclic approximation
(Equation~\eqref{eq:Ji_err_mosttime}), which shows excellent agreement with
the numerically computed values.}
\label{fig:dJz_fun_Jz}
\end{figure}

To further understand the effect of the midplane error, we also plot the
fractional error in $J_z$ as a function of $J_z$ for $z$ offsets of $10, 50,
100\,\pc$ (orange, teal, and green, respectively) in
Figure~\ref{fig:dJz_fun_Jz}. For each orbit, we set the initial position to be $(8,0,0)\,\kpc$ and the initial velocity to be $(-190, 0, v_z)\,\kms$, where we vary $v_z$. For a thin disk-like orbit
($J_z\sim0.5\,\actunit$), even a $10\,\pc$ offset in $z$ is enough to induce a
$\sim25\%$ error in $J_z$. For larger values of $J_z$, the fractional errors
are suppressed, but the induced error can still be large depending on how
great the $z$ offset is. We also plot the epicylic prediction for $\Delta J_z
/ J_z$ from Equation~\eqref{eq:Ji_err_mosttime} as dashed lines for each $z$
offset. We find that the epicyclic approximation matches the numerical
estimate quite well.

@@

\section{LSR Variations} \label{app:lsr}
We consider the variations of the LSR as a function of azimuth at the fiducial
solar circle ($R_{0} = 8.2\,\kpc$) in Figure~\ref{fig:lsr_variations}. At
each azimuth, $\phi$, we take the median velocity in cylindrical coordinates
of all stars within $200\,\pc$ of the position, following
\citet{2018arXiv180610564S}. No best-fit subtraction was performed as in
Figure~\ref{fig:midplane}.
